Ultrafast Electron Microscopes: Design Criteria, Electron Sources, and Column Modeling

### 2013-04-09 inc/photocathode/este/este.tex
more tweaks to the photocathode sections

\ref{fig:este-semicond} depicts the results obtained for GaSb and InSb photocathode materials under pulsed UV (261nm) laser irradiation using the same experimental technique.
Both samples are cut from [100]-oriented polished wafers; the GaSb is undoped and the InSb is moderately $p$-type.
Prior studies \cite{gobeli_photoelectric_1965} indicate that single-photon photoemission should not possible for either semiconductor --- the effective photoemission work functions (about 4.8eV for undoped GaSb and 4.89eV for $p$-type InSb (Fermi level pinned at valence band maximum)) being greater than the 4.75eV photon energy even when the 34meV Schottky barrier suppression due to the applied 8kV/cm DC field is included \cite{dowell_quantum_2009}.
Nonetheless, for both zincblende semiconductors, significant laser-driven emission (more than expected for a single-photon field emission process) is observed with a yield nearly linearly proportional to the $\sim$4ps UV laser pulse energy.

@@

\ref{fig:este-semicond} also shows that the measured Fourier plane electron beam spot size for GaSb is independent of the UV laser pulse energy, as would be expected for the proposed emission process.
More significantly, using $\Delta p_{\smallT} = \sqrt{m k_B T_e}$ (i.e., assuming the emitted electrons have a mass $m$ in the semiconductor), simulation of the experiment with the extended AG model \cite{berger_semi-analytic_2010} indicates that the observed GaSb Fourier plane spot size of 0.45($\pm$0.03)mm (HWe$^{-1}$M) would be due to an electron temperature $T_e \approx$ 360K which is associated with negligible thermionic emission ($\exp[-\Phi/(k_B T_e)] \sim 10^{-15}$).
On the other hand, employing a mass of 0.3$m$ for the electrons in GaSb allows an average electron temperature (over the $\sim$4ps laser pulse duration) of about 1,200K to be extracted using the AG model simulation --- a value much more consistent with the proposed ESTE mechanism and the expected cooling rate by LO phonon emission.
In fact, at the maximum $\sim$0.2nJ incident UV laser pulse energy up to 108 electrons/pulse could be excited into the upper $\Gamma_8$ conduction band, of which one in $\exp[\Phi/(k_B T_e)] \approx$ 20,000 are above the effective 0.99eV work function, thus agreeing with the observed yield of $\sim$10$^3$ electrons/pulse (\ref{fig:este-semicond}).
We also note that the normalized transverse rms emittance $\varepsilon_{\smallT}$ of this UV laser-driven ultrafast GaSb electron source is more than a factor of two less than that expected from a standard Cu photocathode ($\Phi$ = 4.31eV and assuming $\Delta p_{\smallT} = \sqrt{m ( \hbar \omega - \Phi ) / 3 }$, where $\hbar \omega$ is the incident photon energy \cite{dowell_quantum_2009,jensen_emittance_2010}) irradiated at the same 261nm wavelength.

>>>>

\ref{fig:este-semicond} depicts the results obtained for GaSb and InSb photocathode materials under pulsed UV (261nm) laser irradiation using the same experimental technique.
Both samples are cut from [100]-oriented polished wafers; the GaSb is undoped and the InSb is moderately $p$-type.
Prior studies \cite{gobeli_photoelectric_1965} indicate that single-photon photoemission should not possible for either semiconductor --- the effective photoemission work functions (about 4.8eV for undoped GaSb and 4.89eV for $p$-type InSb (Fermi level pinned at valence band maximum)) being greater than the 4.75eV photon energy even when the 34meV Schottky barrier suppression due to the applied 8kV/cm DC field is included \cite{dowell_quantum_2009}.
Nonetheless, for both zincblende semiconductors, significant laser-driven emission (more than expected for a single-photon field emission process) is observed with a yield nearly linearly proportional to the $\sim$4ps UV laser pulse energy.
Note that due to the surprisingly large yield, we repeated the measurements, this time replacing the the scintillator screen with a Faraday cup and repeating the laser energy sequence.
In this way, the yields for these photocathodes were measured directly and absolutely, unlike the results presented in Sections \ref{sec:single_photon} and \ref{sec:two_photon_thermionic}, for which relative measurements were sufficient.

@@

\ref{fig:este-semicond} also shows that the measured Fourier plane electron beam spot size for GaSb is independent of the UV laser pulse energy, as would be expected for the proposed emission process.
More significantly, using $\Delta p_{\smallT} = \sqrt{m k_B T_e}$ (i.e., assuming the emitted electrons have a mass $m$ in the semiconductor), simulation of the experiment with the extended AG model \cite{berger_semi-analytic_2010} indicates that the observed GaSb Fourier plane spot size of 0.45($\pm$0.03)mm (HWe$^{-1}$M) would be due to an electron temperature $T_e \approx$ 360K which is associated with negligible thermionic emission ($\exp[-\Phi/(k_B T_e)] \sim 10^{-15}$).
On the other hand, employing a mass of 0.3$m$ for the electrons in GaSb allows an average electron temperature (over the $\sim$4ps laser pulse duration) of about 1,200K to be extracted using the AG model simulation --- a value much more consistent with the proposed ESTE mechanism and the expected cooling rate by LO phonon emission.
In fact, at the maximum $\sim$0.2nJ incident UV laser pulse energy up to $10^8$ electrons/pulse could be excited into the upper $\Gamma_8$ conduction band, of which one in $\exp[\Phi/(k_B T_e)] \approx$ 20,000 are above the effective 0.99eV work function, thus agreeing with the observed yield of $\sim$10$^3$ electrons/pulse (\ref{fig:este-semicond}).
We also note that the normalized transverse rms emittance $\varepsilon_{\smallT}$ of this UV laser-driven ultrafast GaSb electron source is more than a factor of two less than that expected from a standard Cu photocathode ($\Phi$ = 4.31eV and assuming $\Delta p_{\smallT} = \sqrt{m ( \hbar \omega - \Phi ) / 3 }$, where $\hbar \omega$ is the incident photon energy \cite{dowell_quantum_2009,jensen_emittance_2010}) irradiated at the same 261nm wavelength.

### 2013-04-09 inc/photocathode/future.tex
more tweaks to the photocathode sections

The philosophical argument presented in Section \ref{sec:este} for using the effective mass $m^*$ in the expressions for the transverse momentum variance $\Delta p_{\smallT}$ simply follows from conservation of energy and momentum at the barrier and actually does not depend on the ``excited state'' nature nor the thermionic final emission process.
The earlier experiment employing two-photon assisted thermionic emission (2PTE) from gold (Section \ref{sec:two_photon_thermionic}), the close agreement between the data and the simulation is supported by the fact that the effective mass of a free electron in Au is equal to the rest mass $m$ \cite{johnson_optical_1972}.

For direct single-photon photoemission, we would therefore also expect to require that the expression describing the rms transverse momentum to be written as $\Delta p_{\smallT} = \sqrt{m^* ( \hbar \omega - \Phi ) / 3 }$.
Efforts are currently underway to determine if this is indeed the case using a variety of metal photocathodes and the $\sim$4ps, 261nm laser radiation source.
In an upcoming publication, we will show that the dependence of $\Delta p_{\smallT}$ on the electron effective mass may be extended to simple planar metal photocathodes --- thus potentially allowing for the future development or discovery of a robust, high brightness photocathode material.
The simulation is in good agreement with our direct measurements of $\Delta p_{\smallT}$ for Bi(111) and polycrystalline Ag, Mo, Ta, Be, and W photocathodes.
We will also discuss the effect of laser-induced heatingon $\Delta p_{\smallT}$ of the electron distribution in the photocathode during photoemission, which may limit the attainable pulsed electron source brightness at the desired high peak pulse current densities.

>>>>

The philosophical argument presented in Section \ref{sec:este} for using the effective mass $m^*$ in the expressions for the transverse momentum variance $\Delta p_{\smallT}$ simply follows from conservation of energy and momentum at the barrier and actually does not depend on the ``excited state'' nature nor the thermionic emission process.
For the example two-photon assisted thermionic emission (2PTE) from gold presented in Section \ref{sec:two_photon_thermionic}, the close agreement between the data and the simulation is supported by the fact that the effective mass of a free electron in Au is equal to the rest mass $m$ \cite{johnson_optical_1972}.

For direct single-photon photoemission, we would therefore also expect to require that the expression describing the rms transverse momentum to be written as $\Delta p_{\smallT} = \sqrt{m^* ( \hbar \omega - \Phi ) / 3 }$.
Efforts are currently underway to determine if this is indeed the case using a variety of metal photocathodes and the $\sim$4ps, 261nm laser radiation source.
In an upcoming publication, we will show that the dependence of $\Delta p_{\smallT}$ on the electron effective mass may be extended to simple planar metal photocathodes --- thus potentially allowing for the future development or discovery of a robust, high brightness photocathode material.
For a selection of single-crystal and polycrystalline photocathodes, we will show that the simulation, incorporating $m^*$, is in good agreement with our direct measurements of $\Delta p_{\smallT}$.
We will also discuss the effect of laser-induced heatingon $\Delta p_{\smallT}$ of the electron distribution in the photocathode during photoemission, which may limit the attainable pulsed electron source brightness at the desired high peak pulse current densities.

### 2013-04-09 inc/photocathode/two_photon_thermionic/two_photon_thermionic.tex
more tweaks to the photocathode sections

\begin{figure}
  \centering
  \input{au_plot}
  \caption[Measured Fourier plane electron beam size and yield vs incident laser pulse energy on gold photocathode]{
    Measured Fourier plane electron beam size (HWe$^{-1}$M; squares) and electron yield (circles) as a function of the incident $\sim$200fs, 523nm laser pulse energy for the 300nm-thick gold photocathode.
    The electron yield is proportional to the square of laser pulse energy (solid line) and the dependence of the beam size with the laser pulse energy fits that predicted by a zero free parameter electron heating model (solid line; dashed lines represent $\pm$10\% error in the laser spot size).
    A representative raw Fourier plane beam image is also shown.
  }
  \label{fig:gold-emission}
\end{figure}

@@

The quadratic dependence of the electron yield on pulse energy clearly indicates a two-photon emission process.
This is expected since an excitation energy of two green photons ($\hbar \omega$ = 2.37eV) will be required to overcome the reported $\Phi$ = 4.69eV effective work function of a thick gold film contaminated with adsorbed water \cite{monjushiro_ultraviolet_1991} or ensure that at least the tail of the electron Fermi distribution has sufficient energy to overcome the 5.1($\pm$0.1)eV work function of a clean Au surface \cite{eastman_photoelectric_1970}.
Moreover, when $2\hbar \omega \approx \Phi$, the dependence of the yield on the square of the electron's excess energy above the work function \cite{monjushiro_ultraviolet_1991} ensures that predominantly only the high energy (Boltzmann) tail of the Fermi distribution contributes to the emission.
Consequently, the observed increase in $\Delta p_{\smallT}$ with incident pulse energy must be due to a heating effect; specifically, laser heating of the free electron Fermi gas as two-photon excitation is instantaneous and the $\sim$200fs green laser pulse duration excludes any coupling to the lattice which occurs on the time scale of a few picoseconds \cite{chen_semiclassical_2006}.
The solid line in \ref{fig:gold-emission} is the predicted variation derived using a simple zero free parameter model of this effect.
Using the known optical properties of gold \cite{johnson_optical_1972}, which give a reflectance $R_p$ = 0.53 and an absorption depth of 20nm, and the temperature-dependent heat capacity of a free electron Fermi gas, the model evaluates the average temperature of the laser-heated electron gas that may be two-photon excited assuming that only electrons within a few nanometers of the surface can be emitted, as the mean free path for electrons in Au is $\sim$4nm \cite{seah_quantitative_1979}.
This average electron temperature $T_e$ is then used in the AG model simulation of the experiment (\ref{fig:transverse-measurement}) to determine the expected Fourier plane spot size on the YAG scintillator; that is, using an initial $\Delta p_{\smallT} = \sqrt{m k_B T_e}$.  The dashed lines indicate the expected trends for a $\pm$10\% error in the laser spot size $\Delta x_0$ incident on the gold cathode surface.
The close agreement between the data and the simulation strongly supports our interpretation of the emission mechanism.

>>>>

\begin{figure}
  \centering
  \input{au_plot}
  \caption[Measured Fourier plane electron beam size and yield vs incident laser pulse energy on gold photocathode]{
    Measured Fourier plane electron beam size (HWe$^{-1}$M; squares) and electron yield (circles) as a function of the incident $\sim$200fs, 523nm laser pulse energy for the 300nm-thick gold photocathode.
    The electron yield is proportional to the square of laser pulse energy (solid green line) and the dependence of the beam size with the laser pulse energy fits that predicted by a zero free parameter electron heating model (solid red line; dashed lines represent $\pm$10\% error in the laser spot size).
    A representative raw Fourier plane beam image is also shown.
  }
  \label{fig:gold-emission}
\end{figure}

@@

The quadratic dependence of the electron yield on pulse energy clearly indicates a two-photon emission process.
This is expected since an excitation energy of two green photons ($\hbar \omega$ = 2.37eV) will be required to overcome the reported $\Phi$ = 4.69eV effective work function of a thick gold film contaminated with adsorbed water \cite{monjushiro_ultraviolet_1991} or ensure that at least the tail of the electron Fermi distribution has sufficient energy to overcome the 5.1($\pm$0.1)eV work function of a clean Au surface \cite{eastman_photoelectric_1970}.
Moreover, when $2\hbar \omega \approx \Phi$, the dependence of the yield on the square of the electron's excess energy above the work function \cite{monjushiro_ultraviolet_1991} ensures that predominantly only the high energy (Boltzmann) tail of the Fermi distribution contributes to the emission.
Consequently, the observed increase in $\Delta p_{\smallT}$ with incident pulse energy must be due to a heating effect; specifically, laser heating of the free electron Fermi gas as two-photon excitation is instantaneous and the $\sim$200fs green laser pulse duration excludes any coupling to the lattice which occurs on the time scale of a few picoseconds \cite{chen_semiclassical_2006}.
The solid red line in \ref{fig:gold-emission} is the predicted variation derived using a simple zero free parameter model of this effect.
Using the known optical properties of gold \cite{johnson_optical_1972}, which give a reflectance $R_p$ = 0.53 and an absorption depth of 20nm, and the temperature-dependent heat capacity of a free electron Fermi gas, the model evaluates the average temperature of the laser-heated electron gas that may be two-photon excited assuming that only electrons within a few nanometers of the surface can be emitted, as the mean free path for electrons in Au is $\sim$4nm \cite{seah_quantitative_1979}.
This average electron temperature $T_e$ is then used in the AG model simulation of the experiment (\ref{fig:transverse-measurement}) to determine the expected Fourier plane spot size on the YAG scintillator; that is, using an initial $\Delta p_{\smallT} = \sqrt{m k_B T_e}$.  The dashed red lines indicate the expected trends for a $\pm$10\% error in the laser spot size $\Delta x_0$ incident on the gold cathode surface.
The close agreement between the data and the simulation strongly supports our interpretation of the emission mechanism.
